\section{Threshold Sensitivity}
\label{app:sensitivity}
To establish the robustness of our diagnostic stratification, we conduct a sensitivity analysis evaluating whether the observed predictive degradation and calibration decay are sensitive to the choice of frequency thresholds. We evaluate three alternative threshold configurations: (i) an aggressive setting with thresholds \{10, 50, 250\} training interactions, (ii) a conservative setting with thresholds \{30, 150, 750\} training interactions, and (iii) a dynamic quantile-based bucketing scheme where buckets are defined by equal-quantile KCs. The results, summarized in Table~\ref{tab:sensitivity}, show that while the absolute counts of KCs in each category shift, the diagnostic patterns are broadly consistent, although effect magnitudes vary across threshold definitions. Overall, low-frequency buckets often exhibit higher or more variable calibration errors, although the exact ordering is not strictly monotonic across all threshold settings.

\input{tables/tableA1_sensitivity}
\FloatBarrier

\section{Overall and Bucket-level Performance under Temporal Splits}
\label{app:temporal_performance}
In this section, we present the comprehensive overall performance (Table~\ref{tab:overall_temporal}) and strata-specific performance breakdowns (Table~\ref{tab:bucket_temporal}) under the time-partitioned (temporal) split mode. As shown, deep learning baselines undergo noticeable calibration degradation and predictive decay compared to their population-level learner-based counterparts, highlighting the increased difficulty of curriculum-level forecasting.

\input{tables/tableA_overall_full}
\input{tables/tableA_performance_by_bucket_full}

\clearpage
\onecolumn
\section{Detailed Calibration Results}
\label{app:calibration_performance}
Detailed calibration results by frequency stratum are reported in Table~X.

\input{tables/table_v_calibration_by_bucket_updated}
\clearpage
\twocolumn

\section{Diagnostic Interpretation Guide}
\label{app:interpretation}
To assist researchers and practitioners in applying our diagnostic protocol, we provide Table~\ref{tab:interpretation_guide} which maps common empirical diagnostic patterns observed during evaluation to their theoretical and operational educational interpretation.

\input{tables/table_interpretation_guide}

\section{Bayesian Knowledge Tracing Numerical Instability and Transition}
\label{app:bkt_instability}
During our initial protocol implementation, we instantiated Bayesian Knowledge Tracing (BKT) as the classical baseline utilizing the popular \texttt{pyBKT} library (v1.4.1). However, on all three large-scale datasets, BKT suffered from severe numerical instability during Expectation-Maximization (EM) parameter estimation. Specifically, for sparse and very sparse concept strata, the scarcity of interaction logs triggered a divide-by-zero error in the M-step update equation for the prior parameter, resulting in \texttt{NaN} value assignments. Consequently, the model's predictions degenerated into deterministic binary outputs ($p_{pred} \in \{0.0, \text{NaN}\}$), yielding a flat $0.50$ AUC and extreme Negative Log-Likelihood (NLL) values ranging between $23.7$ and $28.0$. To maintain scientific integrity and standard baseline comparison, we transitioned to Item Response Theory (IRT 1PL / Rasch Model) as the primary classical baseline. The raw prediction CSV files for the failed BKT runs are preserved within the codebase predictions directory for transparency.

