\section{Conclusion}
\label{sec:conclusion}
This paper presented a reproducible diagnostic protocol for sparse-concept and calibration evaluation in Knowledge Tracing, combining train-only KC-frequency stratification with an explicit strict cold-start group, learner-based and temporal splits, per-stratum calibration analysis (ECE and the Brier decomposition, with reliability diagrams), sample-size-aware reliability flags, and an eight-channel leakage and predictive-sanity audit. The protocol is intended as reusable evaluation infrastructure for KT research, in particular for future graph-based and self-supervised approaches whose claimed gains concentrate precisely on sparse KCs, where evaluation is least stable.

The instantiation on three datasets illustrates what the protocol surfaces that aggregate metrics do not. While population-level metrics remained competitive, calibration degraded as KC training evidence decreased. For example, on ASSISTments 2012, SimpleKT's ECE increased consistently from $0.113$ on dense concepts to $0.158$ on medium concepts and $0.225$ on sparse concepts. The latter figure should be interpreted with caution given its Limited reliability flag ($N = 403$). The heterogeneity of calibration and performance trends across datasets further underscores why a standardized, dataset-agnostic protocol is needed: sparse-concept behavior cannot be extrapolated from one dataset to another. Finally, the temporal-split audit provided an empirical demonstration of the protocol's hygiene value: the predictive-sanity channel caught a prediction--label alignment artifact that would otherwise have been misread as genuine cold-start failure, allowing researchers to distinguish concept-level cold-start difficulty from pipeline errors.

Future work can extend this diagnostic foundation by evaluating across multiple temporal seeds or cutoffs to capture variance, by incorporating a classical baseline that exploits item difficulty for unseen learners under learner-based splits, and by applying the protocol to diagnose graph-augmented and self-supervised KT architectures.
