\section{Threshold Sensitivity Protocol}
\label{app:sensitivity}
Threshold sensitivity analysis is implemented in the released artifact as an optional robustness check. The present manuscript does not use threshold-sensitivity results as primary evidence because the main diagnostic conclusions are based on the pre-registered KC-frequency thresholds used in Tables 5, 6, 9, and 10. The artifact evaluates three alternative threshold configurations: (i) an aggressive setting with thresholds \{10, 50, 250\} training interactions, (ii) a conservative setting with thresholds \{30, 150, 750\} training interactions, and (iii) a dynamic quantile-based bucketing scheme where buckets are defined by equal-quantile KCs. These checks are intended to help future users evaluate whether their conclusions are sensitive to the chosen KC-frequency thresholds.

\FloatBarrier

\section{Overall and Bucket-level Performance under Temporal Splits}
\label{app:temporal_performance}
In this section, we present the detailed overall performance (Table~\ref{tab:temporal_overall}) and strata-specific performance breakdowns (Table~\ref{tab:temporal_strata}) under the time-partitioned (temporal) split mode. As shown, deep learning baselines undergo noticeable predictive decay compared to their population-level learner-based counterparts, highlighting the increased difficulty of curriculum-level forecasting. Calibration degradation under temporal splits is reported separately in Appendix~\ref{app:calibration_performance}, Table~\ref{tab:calibration_temporal}.

\input{tables/tableA_overall_full}
\input{tables/tableA_performance_by_bucket_full}

\clearpage
\section{Detailed Calibration Results}
\label{app:calibration_performance}
Detailed calibration results by frequency stratum for the learner-based split are reported in Table~\ref{tab:calibration_learner}. The corresponding temporal calibration breakdowns, computed after the label-alignment correction documented in Appendix~\ref{app:alignment} and used as the numerical basis for Figure~\ref{fig:reliability_junyi_temporal}, are reported in Table~\ref{tab:calibration_temporal}.

\input{tables/table7_calibration_by_bucket}
\input{tables/table8_temporal_calibration}
\clearpage

\section{Diagnostic Interpretation Guide}
\label{app:interpretation}
To assist researchers and practitioners in applying our diagnostic protocol, we provide Table~\ref{tab:interpretation_guide} which maps common empirical diagnostic patterns observed during evaluation to their theoretical and operational educational interpretation.

\input{tables/table_interpretation_guide}

\section[Bayesian Knowledge Tracing Numerical Instability and Transition]{Bayesian Knowledge Tracing Numerical Instability and Transition}
\label{app:bkt_instability}
During our initial protocol implementation, we instantiated Bayesian Knowledge Tracing (BKT) as the classical baseline utilizing the popular \texttt{pyBKT} library (v1.4.1). However, on all three large-scale datasets, BKT suffered from severe numerical instability during Expectation-Maximization (EM) parameter estimation. Specifically, for sparse and very sparse concept strata, the scarcity of interaction logs triggered a divide-by-zero error in the M-step update equation for the prior parameter, resulting in \texttt{NaN} value assignments. Consequently, the model's predictions degenerated into deterministic binary outputs ($p_{pred} \in \{0.0, \text{NaN}\}$), yielding a flat $0.50$ AUC and extreme Negative Log-Likelihood (NLL) values ranging between $23.7$ and $28.0$. To maintain scientific integrity and standard baseline comparison, we transitioned to Item Response Theory (IRT 1PL / Rasch Model) as the primary classical baseline. The raw prediction CSV files for the failed BKT runs are preserved within the codebase predictions directory for transparency.

\section{Prediction--Label Alignment Audit for Temporal Splits}
\label{app:alignment}
During the temporal-split audit, we identified a prediction--label alignment issue in the temporal evaluation pipeline. The issue occurred when exported model predictions were matched back to temporal test instances after sequence construction and cohort filtering. In the affected code path, prediction rows and correctness labels could be aligned by transient row order rather than by a stable instance identifier, which made some exported $p_{pred}$ values correspond to incorrect $y_{true}$ entries.

The main symptom was not limited to strict cold-start KCs. Deep KT baselines produced near-random AUC even on warm temporal cohorts, where KCs had sufficient training-fold evidence. This behavior was a red flag because a static IRT baseline still achieved meaningful AUC on the same warm cohorts, whereas DKT and SimpleKT collapsed to approximately random ranking.

We detected the issue using two sanity checks. First, we compared warm-cohort AUC against the IRT reference baseline; near-random deep KT AUC on warm KCs was treated as an alignment warning rather than as evidence of genuine temporal generalization failure. Second, we computed prediction--label correlation on warm-cohort slices and verified whether $\mathbb{E}[p_{pred}\mid y=1]$ exceeded $\mathbb{E}[p_{pred}\mid y=0]$. The correction preserves a stable test-instance identity through temporal splitting, sequence construction, prediction export, cohort filtering, and metric computation. Metrics are computed only after pairing predictions and labels by this stable identity rather than by transient row position.

After applying the correction consistently across datasets, warm-cohort AUC for the affected deep KT baselines recovered above the sanity threshold, and prediction--label correlations became positive. These checks were then used before re-running the temporal diagnostics reported in the paper. This audit reinforces the central message of the protocol: temporal cold-start results should be accompanied by warm-cohort and prediction--label sanity checks to distinguish genuine concept-level cold-start difficulty from alignment artifacts.
