\section{Discussion and Limitations}
\label{sec:discussion}
In this section, we discuss the broader implications of our diagnostic protocol for educational technology research and practice. We then define the scope boundaries of our work and address the primary limitations of the current study.

\subsection{Implications}
The protocol encourages KT evaluation to move beyond aggregate AUC in three concrete ways:
\begin{itemize}
    \item \textbf{For researchers:} Per-stratum reporting with reliability flags (Section 3.5) makes claims about sparse and cold-start KCs auditable. A reported gain on low-frequency concepts is interpretable only together with its stratum sample size, and the protocol makes that context mandatory rather than optional.
    \item \textbf{For developers of educational systems:} The calibration diagnostics speak directly to deployment practice. When predicted probabilities drive remediation, practice, or advancement decisions through fixed thresholds, a model whose ECE roughly doubles from dense to sparse strata (a Limited-reliability stratum)---as observed for SimpleKT on ASSISTments 2012 under our experimental conditions---may distort threshold-based intervention decisions, especially in low-frequency strata where instructional support is most sensitive to probability reliability. Threshold-based intervention logic should therefore be validated per stratum, not on population-level calibration alone.
    \item \textbf{For the reviewing and benchmarking community:} The audit demonstrates its practical value. Channel L8 caught a prediction--label misalignment that aggregate metrics would have misreported as a scientific finding about temporal cold-start difficulty (Section~\ref{sec:threats}). Evaluation hygiene of this kind is inexpensive relative to the cost of a corrupted conclusion entering the literature.
\end{itemize}

\subsection{Scope Boundaries}
This paper is intentionally limited to evaluation diagnostics. It does not introduce a new KT model, a self-supervised learning (SSL) objective, a graph neural network (GNN), a graph augmentation strategy, a learning path recommender (LPR), or a distillation method. Those directions require separate methodological papers and should be evaluated using the protocol introduced here. A complementary line of work addresses leakage control in concept-graph construction for graph-augmented KT; such leakage-control protocols and the diagnostics proposed here are complementary and can be combined in future structure-aware KT studies.

\subsection{Threats to Validity and Limitations}
\label{sec:threats}
We analyze the threats to the validity of our diagnostic protocol across standard dimensions:
\begin{itemize}
    \item \textbf{Internal Validity:} Risks to internal validity include potential preprocessing bugs, split leakage, sequence--label alignment errors, and implementation-specific training differences across the baselines. We mitigate these risks through a standardized interaction data schema, rigid validation-fold early stopping, and our automated leakage audit checklist. Additionally, we note that our classical baseline (IRT) under learner-based splits acts as a base-rate reference due to the lack of ability estimates for unseen learners; developing a classical baseline that effectively exploits item difficulty for unseen learners would strengthen future comparative work.
    \item \textbf{Temporal Alignment Validity:} Sequence modeling in KT is notoriously prone to ``silent bugs'' such as off-by-one prediction--label misalignments. Our protocol actively defends against this via the warm-cohort predictive sanity audit (L8). During validation, this automated check successfully flagged an alignment discrepancy in the temporal evaluation pipeline that otherwise would have corrupted the evaluation.
    \item \textbf{Construct Validity:} Construct validity refers to whether our metrics accurately capture latent learner mastery and concept calibration. In KT, KC definitions depend heavily on dataset annotation quality, Q-matrix provenance, and how multi-skill items are processed. For example, expert-tagged KCs in ASSISTments 2012 and Junyi Academy may introduce varying granularities, which directly impact strata boundaries.
    \item \textbf{External Validity:} The pronounced heterogeneity observed across datasets---such as the clear calibration degradation gradient in ASSISTments 2012, the inverted calibration patterns in XES3G5M, and the entirely empty very-sparse stratum in Junyi Academy---supports the necessity of the protocol. Models cannot be assumed to behave consistently across different sparsity distributions, making dataset-local validation essential.
    \item \textbf{Statistical Conclusion Validity:} very sparse buckets can contain few KCs and few evaluation events (as few as 3 to 10 events per seed in our instantiation), which inevitably produces metric instability and high variance; the reliability flags of Section 3.5 exist precisely to keep such estimates from being over-interpreted. Two further limitations belong under this heading. First, the temporal-split diagnostics are computed from a single corrected evaluation run at a fixed chronological cutoff (Section 4.1); they therefore characterize a fixed-cutoff stress test and do not estimate variance across temporal cut-points or neural initializations. Extending the temporal diagnostics to multiple seeds and multiple cutoffs is a direct avenue for future work. Second, because the Brier decomposition is computed from binned probabilities (Section 3.4), its components are empirical approximations and need not sum exactly to the directly computed Brier score; decomposition values should be read as diagnostic indicators rather than exact accounting identities.
\end{itemize}
