% -----------------------------------------------------------------------------
% JEDM LaTeX Format for P0 Paper - ANONYMOUS PEER REVIEW
% Title: Reproducible Sparse-Concept and Calibration Diagnostics for Knowledge Tracing
% -----------------------------------------------------------------------------
\documentclass[12pt,a4paper]{jedm}

% --- JEDM Formatting ---
\usepackage[left=2.75cm, right=2.75cm, top=2.25cm, bottom=2.25cm]{geometry}
\usepackage{mathptmx} % Times Roman for body
\usepackage[scaled=0.92]{helvet} % Helvetica for sans-serif
\usepackage{titlesec} % For heading fonts
\titleformat*{\section}{\Large\bfseries\sffamily}
\titleformat*{\subsection}{\large\bfseries\sffamily}
\titleformat*{\subsubsection}{\normalsize\bfseries\sffamily}

\usepackage{caption}
\captionsetup[table]{position=top, skip=5pt}
\captionsetup[figure]{position=bottom, skip=5pt}

% Packages
\usepackage{cite} % For numerical citations [1], [2] (Standard/Robust)
% If you prefer JEDM's official author-year citations, comment the line above and uncomment the line below:
% \usepackage[round,authoryear]{natbib}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{array}
\usepackage{xcolor}
\usepackage{url}
\usepackage{listings}
\usepackage[hidelinks]{hyperref}
\usepackage{subcaption}
\usepackage{placeins}
\usepackage{float}
\usepackage{tabularx}
\usepackage{makecell}
\usepackage{adjustbox}

% Float control parameters to prevent whitespace gaps
\renewcommand{\topfraction}{0.95}
\renewcommand{\bottomfraction}{0.95}
\renewcommand{\textfraction}{0.05}
\renewcommand{\floatpagefraction}{0.95}

% Code listing style
\lstdefinestyle{p0style}{
  basicstyle=\ttfamily\footnotesize,
  breaklines=true,
  frame=single,
  columns=fullflexible,
  keepspaces=true,
  showstringspaces=false,
  tabsize=2
}
\lstset{style=p0style}

% Macros
\newcommand{\pzero}{P0}
\newcommand{\kt}{Knowledge Tracing}
\newcommand{\kc}{Knowledge Component}
\newcommand{\kcs}{Knowledge Components}
\newcommand{\ece}{\mathrm{ECE}}
\newcommand{\brier}{\mathrm{Brier}}
\newcommand{\unc}{\mathrm{UNC}}
\newcommand{\rel}{\mathrm{REL}}
\newcommand{\res}{\mathrm{RES}}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}

% For compatibility with tables that use tableorg
\let\tableorg\table
\let\endtableorg\endtable

\begin{document}

\title{\sffamily\bfseries Reproducible Sparse-Concept and Calibration Diagnostics for Knowledge Tracing\\
{\large A Protocol and Diagnostic Study}}

\author{
\sffamily Anonymous Authors\\[8pt]
\normalsize \textit{Anonymous Affiliation}\\
\normalsize Email: anonymous@example.com
}
\date{}

\maketitle

\begin{abstract}
Knowledge Tracing (KT) models are commonly evaluated using aggregate predictive metrics such as AUC and accuracy. However, overall metrics may hide model behavior on sparse or limited cold‑start Knowledge Components (KCs), where reliable probability estimates are crucial for educational decision support. This paper presents a reproducible diagnostic protocol for sparse‑concept and calibration evaluation in KT. The protocol combines learner‑based and temporal splits, train‑only KC‑frequency stratification, limited cold‑start analysis, calibration metrics (ECE, Brier decomposition), reliability diagrams per stratum, and an eight‑channel leakage and predictive sanity audit checklist. We instantiate the protocol on ASSISTments 2012, Junyi Academy, and XES3G5M with baselines IRT, DKT, and SimpleKT. Under our experimental conditions, on datasets with populated sparse strata, calibration error tends to increase from dense to sparse concepts, while aggregate AUC may obscure stratum-level reliability differences. We prepare scripts and report templates for release upon acceptance to support reproducible sparse‑concept and calibration diagnostics for future KT studies.
\end{abstract}

\vspace{10pt}
\noindent\textbf{Keywords:} Knowledge Tracing; Educational Data Mining; Calibration; Sparse Concepts; Cold-start Knowledge Components; Reproducibility

\input{sections/01_introduction}
\input{sections/02_related_work}
\input{sections/03_protocol}
\input{sections/04_experiments}
\input{sections/05_discussion_limitations}
\input{sections/06_conclusion}

\section*{Artifact Availability \& Reproducibility Checklist}
The code and reporting templates will be made publicly available upon acceptance. Consistent with the design principles of Section~\ref{sec:protocol}, the artifact includes the following standardized components:
\begin{itemize}
  \item \textbf{Dataset Preprocessing:} Scripts that normalize raw data into the schema defined in Section~\ref{sec:protocol}.
  \item \textbf{Split Construction:} Reproducible partition logs and multi‑seed split constructors.
  \item \textbf{Prediction‑level Exports:} CSV files containing baseline predicted probabilities and ground‑truth labels.
  \item \textbf{KC‑Strata Assignment:} Code mapping training‑fold frequencies to strata buckets.
  \item \textbf{Calibration Verification:} Scripts computing Expected Calibration Error (ECE).
  \item \textbf{Brier Decomposition:} Implementation of Uncertainty, Reliability, and Resolution equations.
  \item \textbf{Reliability Diagrams:} Plotting utilities for multi‑strata calibration curves.
  \item \textbf{Leakage \& Predictive Sanity Audit:} Automated logs verifying channels L1--L8.
  \item \textbf{LaTeX Table Exporters:} Python exporters generating book‑tabs tables.
  \item \textbf{One‑Command Reproduction:} Unified script ensuring machine‑independent reproducibility.
\end{itemize}

\appendix
\gdef\thesection{Appendix \Alph{section}}
\input{appendix/appendix_a_sensitivity}

\section*{Acknowledgements}
Omitted for double-blind review.

\section*{Declaration of Generative AI Software Tools in the Writing Process}
During the preparation of this work, the authors used ChatGPT and Google Antigravity to support language polishing, LaTeX formatting, consistency checking, and preparation of reproducibility prompts in the Abstract, Introduction, Methods, Results, Discussion, and Appendix sections. These tools were not used to generate, fabricate, or alter experimental results. After using these tools, the authors reviewed and edited the content as needed and take full responsibility for the content of the publication.

\input{references.bbl} % Pre-compiled numerical citations (Standard/Robust)
% If you prefer compiling using JEDM's acmtrans BibTeX style, comment the line above and uncomment the two lines below:
% \bibliographystyle{acmtrans}
% \bibliography{references}

\end{document}
