% -----------------------------------------------------------------------------
% Springer Nature Format for Applied Intelligence
% Title: Reproducible Sparse-Concept and Calibration Diagnostics for Knowledge Tracing
% -----------------------------------------------------------------------------
\documentclass[sn-mathphys]{sn-jnl}
\def\iscompact{1}

% Packages

\usepackage{graphicx}
\usepackage{adjustbox}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{placeins}
\usepackage{makecell}
\usepackage[title]{appendix}
\usepackage{xcolor}
\usepackage{textcomp}
\usepackage{manyfoot}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{listings}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{url}
\usepackage{placeins}
% \usepackage[hidelinks]{hyperref}

% Code listing style
\lstdefinestyle{p0style}{
  basicstyle=\ttfamily\footnotesize,
  breaklines=true,
  frame=single,
  columns=fullflexible,
  keepspaces=true,
  showstringspaces=false,
  tabsize=2
}
\lstset{style=p0style}

% Macros
\newcommand{\pzero}{P0}
\newcommand{\kt}{Knowledge Tracing}
\newcommand{\kc}{Knowledge Component}
\newcommand{\kcs}{Knowledge Components}
\newcommand{\ece}{\mathrm{ECE}}
\newcommand{\brier}{\mathrm{Brier}}
\newcommand{\unc}{\mathrm{UNC}}
\newcommand{\rel}{\mathrm{REL}}
\newcommand{\res}{\mathrm{RES}}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}

\begin{document}

\title[Reproducible Sparse-Concept and Calibration Diagnostics]{Reproducible Sparse-Concept and Calibration Diagnostics for Knowledge Tracing: A Protocol and Diagnostic Study}

\author[1]{\fnm{Nguyen} \sur{Khanh Trinh}}\email{trinhnk@utehy.edu.vn}
\author[1]{\fnm{Dao} \sur{Minh Tuan}}\email{tuanymc@utehy.edu.vn}
\author[1]{\fnm{Nguyen} \sur{Tien Duong}}\email{duongnt@utehy.edu.vn}
\author*[1]{\fnm{Nguyen} \sur{Van Hau}}\email{haunv@utehy.edu.vn}
\author[2]{\fnm{Nguyen} \sur{Chi Thanh}}\email{thanhnc@ioit.ai.vn}

\affil*[1]{\orgname{Hung Yen University of Technology and Education}, \orgaddress{\city{Hung Yen}, \country{Vietnam}}}
\affil[2]{\orgname{Academy of Military Science and Technology}, \orgaddress{\city{Ha Noi}, \country{Vietnam}}}

\abstract{Knowledge Tracing (KT) models are commonly evaluated using aggregate predictive metrics such as AUC and accuracy. However, overall metrics may hide model behavior on sparse or limited cold-start Knowledge Components (KCs), where reliable probability estimates are crucial for educational decision support. This paper presents a reproducible diagnostic protocol for sparse-concept and calibration evaluation in KT. The protocol combines learner-based and temporal splits, train-only KC-frequency stratification, limited cold-start analysis, calibration metrics (ECE, Brier decomposition), reliability diagrams per stratum, and an eight-channel leakage and predictive sanity audit checklist. We instantiate the protocol on ASSISTments 2012, Junyi Academy, and XES3G5M with baselines IRT, DKT, and SimpleKT. Under our experimental conditions, on datasets with populated sparse strata, calibration error tends to increase from dense to sparse concepts, while aggregate AUC may obscure stratum-level reliability differences. We prepare scripts and report templates for release upon acceptance to support reproducible sparse-concept and calibration diagnostics for future KT studies.}

\keywords{Knowledge Tracing, Educational Data Mining, Calibration, Sparse Concepts, Reproducibility, Intelligent Systems}

\maketitle

\input{sections/01_introduction}
\input{sections/02_related_work}
\input{sections/03_protocol}
\input{sections/04_experiments}
\input{sections/05_discussion_limitations}
\input{sections/06_conclusion}

\backmatter

\section*{Declarations}

\begin{itemize}
    \item \textbf{Funding:} The authors received no specific funding for this work.
    \item \textbf{Competing interests:} The authors declare no competing interests.
    \item \textbf{Ethics approval:} Not applicable.
    \item \textbf{Consent to participate:} Not applicable.
    \item \textbf{Consent for publication:} Not applicable.
    \item \textbf{Data availability:} The datasets analyzed during the current study are available in public repositories (ASSISTments 2012, Junyi Academy, XES3G5M).
    \item \textbf{Code availability:} The code, protocol checklist, and reporting templates are publicly available on the \href{https://github.com/trinhnkt/P0}{GitHub repository}.
    \item \textbf{Authors' contributions:} Nguyen Khanh Trinh: conceptualization, methodology, implementation, experimental analysis, data curation, and writing---original draft. Dao Minh Tuan: supervision, research design, validation, and writing---review and editing. Nguyen Tien Duong: dataset preparation, reproducibility checking, result verification, and manuscript editing. Nguyen Van Hau and Nguyen Chi Thanh: critical review, validation, and academic consultation. All authors read and approved the final manuscript.
    \item \textbf{Use of generative AI tools:} Generative AI tools were used to support language polishing, LaTeX formatting, consistency checking, and preparation of reproducibility prompts. These tools were not used to generate or alter experimental results. All scientific claims, experimental design decisions, numerical results, and final interpretations were checked, verified, and approved by the authors.
    \item \textbf{ORCID:} Nguyen Khanh Trinh (0009-0004-3739-0484), Dao Minh Tuan (0009-0009-1641-6813), Nguyen Tien Duong (0009-0007-1104-7129), Nguyen Van Hau (0000-0002-3256-5626), Nguyen Chi Thanh (0000-0003-4335-7002).
\end{itemize}

\clearpage
\begin{appendices}
\input{appendix/appendix_a_sensitivity}
\end{appendices}


\input{references.bbl}

\end{document}
