% -----------------------------------------------------------------------------
% Springer Nature Format for Applied Intelligence - ANONYMOUS PEER REVIEW
% Title: Reproducible Sparse-Concept and Calibration Diagnostics for Knowledge Tracing
% -----------------------------------------------------------------------------
\documentclass[sn-mathphys]{sn-jnl}

% Packages
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage[title]{appendix}
\usepackage{xcolor}
\usepackage{textcomp}
\usepackage{manyfoot}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{listings}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{adjustbox}
\usepackage{url}
\usepackage{makecell}
\usepackage{placeins}
\usepackage[hidelinks]{hyperref}

% Code listing style
\lstdefinestyle{p0style}{
  basicstyle=\ttfamily\footnotesize,
  breaklines=true,
  frame=single,
  columns=fullflexible,
  keepspaces=true,
  showstringspaces=false,
  tabsize=2
}
\lstset{style=p0style}

% Macros
\newcommand{\pzero}{P0}
\newcommand{\kt}{Knowledge Tracing}
\newcommand{\kc}{Knowledge Component}
\newcommand{\kcs}{Knowledge Components}
\newcommand{\ece}{\mathrm{ECE}}
\newcommand{\brier}{\mathrm{Brier}}
\newcommand{\unc}{\mathrm{UNC}}
\newcommand{\rel}{\mathrm{REL}}
\newcommand{\res}{\mathrm{RES}}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}

\begin{document}

\title[Reproducible Sparse-Concept and Calibration Diagnostics]{Reproducible Sparse-Concept and Calibration Diagnostics for Knowledge Tracing: A Protocol and Diagnostic Study}

\author*[1]{\fnm{Anonymous} \sur{Author(s)}}\email{anonymous@example.com}

\affil*[1]{\orgname{Anonymous Institution}, \orgaddress{\city{Anonymous City}, \country{Anonymous Country}}}

\abstract{Knowledge Tracing (KT) models are commonly evaluated using aggregate predictive metrics such as AUC and accuracy. However, overall metrics may hide model behavior on sparse or limited cold-start Knowledge Components (KCs), where reliable probability estimates are crucial for educational decision support. This paper presents a reproducible diagnostic protocol for sparse-concept and calibration evaluation in KT. The protocol combines learner-based and temporal splits, train-only KC-frequency stratification, limited cold-start analysis, calibration metrics (ECE, Brier decomposition), reliability diagrams per stratum, and an eight-channel leakage and predictive sanity audit checklist. We instantiate the protocol on ASSISTments 2012, Junyi Academy, and XES3G5M with baselines IRT, DKT, and SimpleKT. Under our experimental conditions, on datasets with populated sparse strata, calibration error tends to increase from dense to sparse concepts, while aggregate AUC may obscure stratum-level reliability differences. We prepare scripts and report templates for release upon acceptance to support reproducible sparse-concept and calibration diagnostics for future KT studies.}

\keywords{Knowledge Tracing, Educational Data Mining, Calibration, Sparse Concepts, Reproducibility, Intelligent Systems}

\maketitle

\input{sections/01_introduction}
\input{sections/02_related_work}
\input{sections/03_protocol}
\input{sections/04_experiments}
\input{sections/05_discussion_limitations}
\input{sections/06_conclusion}

\backmatter

\section*{Declarations}

\begin{itemize}
    \item \textbf{Funding:} The authors received no specific funding for this work.
    \item \textbf{Competing interests:} The authors declare no competing interests.
    \item \textbf{Ethics approval:} Not applicable.
    \item \textbf{Consent to participate:} Not applicable.
    \item \textbf{Consent for publication:} Not applicable.
    \item \textbf{Data availability:} The datasets analyzed during the current study are available in public repositories (ASSISTments 2012, Junyi Academy, XES3G5M).
    \item \textbf{Code availability:} The code and reporting templates will be made publicly available upon acceptance.
    \item \textbf{Authors' contributions:} Anonymous for peer review.
    \item \textbf{Use of generative AI tools:} During the preparation of this work, the authors used ChatGPT and Google Antigravity to support language polishing, LaTeX formatting, consistency checking, and preparation of reproducibility prompts in the Abstract, Introduction, Methods, Results, Discussion, and Appendix sections. These tools were not used to generate, fabricate, or alter experimental results. After using these tools, the authors reviewed and edited the content as needed and take full responsibility for the content of the publication.
\end{itemize}

\clearpage
\begin{appendices}
\input{appendix/appendix_a_sensitivity}
\end{appendices}

\input{references.bbl}
\end{document}
