% -----------------------------------------------------------------------------
% IEEE Conference LaTeX Template for P0 Paper
% Title: Reproducible Sparse-Concept and Calibration Diagnostics for Knowledge Tracing
% Recommended compiler on Overleaf: pdfLaTeX
% -----------------------------------------------------------------------------
\documentclass[conference]{IEEEtran}

\IEEEoverridecommandlockouts

% ----- Packages -----
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{array}
\usepackage{xcolor}
\usepackage{url}
\usepackage{listings}
\usepackage[hidelinks]{hyperref}
\usepackage{subcaption}
\usepackage{placeins}

% ----- Code listing style -----
\lstdefinestyle{p0style}{
  basicstyle=\ttfamily\footnotesize,
  breaklines=true,
  frame=single,
  columns=fullflexible,
  keepspaces=true,
  showstringspaces=false,
  tabsize=2
}
\lstset{style=p0style}

% ----- Macros -----
\newcommand{\pzero}{P0}
\newcommand{\kt}{Knowledge Tracing}
\newcommand{\kc}{Knowledge Component}
\newcommand{\kcs}{Knowledge Components}
\newcommand{\ece}{\mathrm{ECE}}
\newcommand{\brier}{\mathrm{Brier}}
\newcommand{\unc}{\mathrm{UNC}}
\newcommand{\rel}{\mathrm{REL}}
\newcommand{\res}{\mathrm{RES}}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}

\begin{document}

\title{Reproducible Sparse-Concept and Calibration Diagnostics for Knowledge Tracing\\
{\large A Protocol and Diagnostic Study}}

\author{\IEEEauthorblockN{
Khanh-Trinh Nguyen\IEEEauthorrefmark{1},
Van-Hau Nguyen\IEEEauthorrefmark{1},
Chi-Thanh Nguyen\IEEEauthorrefmark{2},
Tien-Duong Nguyen\IEEEauthorrefmark{1}, and
Minh-Tuan Dao\IEEEauthorrefmark{1}
}
\IEEEauthorblockA{
\IEEEauthorrefmark{1}Hung Yen University of Technology and Education, Hung Yen, Viet Nam\\
Email: \{trinhnkt, hau666, nduongnt, tuan.ymc\}@gmail.com
}
\IEEEauthorblockA{
\IEEEauthorrefmark{2}Academy of Military Science and Technology, Ha Noi, Viet Nam\\
Email: thanhnc80@gmail.com
}
}

\maketitle

\begin{abstract}
Knowledge Tracing (KT) models are commonly evaluated using aggregate predictive metrics such as AUC and accuracy. However, such overall metrics may obscure model behavior on sparse or limited cold-start knowledge components (KCs), where reliable probability estimates are especially important for educational decision support. This paper presents a reproducible diagnostic protocol for sparse-concept and calibration evaluation in KT. The protocol combines learner-based and temporal splits, train-only KC-frequency stratification, limited cold-start concept analysis, calibration metrics including Expected Calibration Error and Brier decomposition, reliability diagrams per stratum, and a seven-channel leakage-control audit checklist. We instantiate the protocol on ASSISTments 2012, Junyi Academy, and XES3G5M with common KT baselines including BKT, DKT, and SimpleKT. Under our experimental conditions, we report how calibration profiles differ across KC-frequency strata and how model rankings under overall AUC may differ from rankings under sparse-KC calibration metrics. We release scripts and report templates to support reproducible sparse-concept and calibration diagnostics for future KT studies.
\end{abstract}

\begin{IEEEkeywords}
Knowledge Tracing, Educational Data Mining, Calibration, Expected Calibration Error, Brier Score, Sparse Concepts, Cold-start Knowledge Components, Reproducibility
\end{IEEEkeywords}

\input{sections/01_introduction}
\input{sections/02_related_work}
\input{sections/03_protocol}
\input{sections/04_experiments}
\input{sections/05_discussion_limitations}
\input{sections/06_conclusion}

\section*{Artifact Availability \& Reproducibility Checklist}
The anonymized artifact repository will be provided during peer review. The public repository will be released upon acceptance. Under our protocol's design principles, the artifact is prepared for release with the following standardized components:
\begin{itemize}
    \item \textbf{Dataset Preprocessing:} Standardized scripts that normalize raw data into the schema specified in Section~\ref{sec:protocol}.
    \item \textbf{Split Construction:} Reproducible partitioning logs and multi-seed split constructors.
    \item \textbf{Prediction-level Exports:} Standardized CSV outputs containing baseline predicted probabilities and ground-truth labels.
    \item \textbf{KC-Strata Assignment:} Scripts mapping training-fold frequencies to strata buckets.
    \item \textbf{Calibration Verification:} Verification scripts for Expected Calibration Error (ECE).
    \item \textbf{Brier Decomposition:} Implementation of the Uncertainty, Reliability, and Resolution decomposition equations.
    \item \textbf{Reliability Diagrams:} Plotting scripts for generating multi-strata calibration curves.
    \item \textbf{Leakage Audit:} Completed automated logs verifying channels L1--L7.
    \item \textbf{LaTeX Table Exporters:} End-to-end Python exporters compiling metrics directly into standard booktabs LaTeX files.
    \item \textbf{One-Command Reproduction:} A unified execution script ensuring machine-independent reproducibility.
\end{itemize}

\appendices
\input{appendix/appendix_a_sensitivity}

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}
