\section{Experiments}
\label{sec:experiments}

\subsection{Datasets}
The experimental setting uses ASSISTments 2012, Junyi Academy, and XES3G5M. Table~\ref{tab:datasets} reports dataset statistics after preprocessing.

\input{tables/table1_dataset_stats}

\subsection{Baselines}
The baseline set includes BKT, DKT, and SimpleKT. Hyperparameters follow benchmark defaults. Model selection uses validation data only.

\subsection{Overall Baseline Diagnostics}
Table~\ref{tab:overall} reports aggregate metrics. These metrics are used to verify that the pipeline runs correctly; they are not the sole basis for conclusions.

\input{tables/table3_overall_results}

\subsection{Sparse-concept Diagnostics}
Table~\ref{tab:strata_performance} reports AUC, accuracy, and NLL by KC-frequency stratum. The expected diagnostic question is whether aggregate results hide degradation on sparse and very sparse KCs.

\input{tables/table4_metric_per_bucket}

\subsection{Calibration by KC Stratum}
Table~\ref{tab:calibration} reports ECE and Brier decomposition by stratum. Reliability diagrams are generated for selected model--dataset pairs in the artifact folder.

\input{tables/table5_calibration_per_bucket}

\subsection{Limited Cold-start Concept Diagnostics}
We define limited cold-start KCs using a pre-registered threshold $k$ such as $k=10$. Table~\ref{tab:coldstart} compares cold-start and warm KC groups.

\input{tables/table6_cold_start_results}

\subsection{Threshold Sensitivity}
To ensure that conclusions do not depend on a single bucket definition, we repeat the diagnostic analysis with alternative thresholds. Results are reported in Appendix~\ref{app:sensitivity}.
