% -----------------------------------------------------------------------------
% JEDM LaTeX Format
% Title: Reproducible Sparse-Concept and Calibration Diagnostics for Knowledge Tracing
% -----------------------------------------------------------------------------
\documentclass[12pt,a4paper]{jedm}

% --- JEDM Formatting ---
\usepackage[left=2.75cm, right=2.75cm, top=2.25cm, bottom=2.25cm]{geometry}
\usepackage{mathptmx} % Times Roman for body
\usepackage[scaled=0.92]{helvet} % Helvetica for sans-serif
\usepackage{titlesec} % For heading fonts
\titleformat*{\section}{\Large\bfseries\sffamily}
\titleformat*{\subsection}{\large\bfseries\sffamily}
\titleformat*{\subsubsection}{\normalsize\bfseries\sffamily}

\usepackage{caption}
\captionsetup[table]{position=top, skip=5pt}
\captionsetup[figure]{position=bottom, skip=5pt}

% Packages
% \usepackage{cite} % For numerical citations [1], [2] (Standard/Robust)
% If you prefer JEDM's official author-year citations, comment the line above and uncomment the line below:
\usepackage[round,authoryear]{natbib}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{array}
\usepackage{xcolor}
\usepackage{url}
\usepackage{listings}
\usepackage[hidelinks]{hyperref}
\usepackage{subcaption}
\usepackage{placeins}
\usepackage{float}
\usepackage{tabularx}
\usepackage{makecell}
\usepackage{adjustbox}

% Float control parameters to prevent whitespace gaps
\renewcommand{\topfraction}{0.95}
\renewcommand{\bottomfraction}{0.95}
\renewcommand{\textfraction}{0.05}
\renewcommand{\floatpagefraction}{0.95}

% Code listing style
\lstdefinestyle{codestyle}{
  backgroundcolor=\color{backcolour},   
  commentstyle=\color{codegreen},
  keywordstyle=\color{magenta},
  numberstyle=\tiny\color{codegray},
  stringstyle=\color{codepurple},
  basicstyle=\ttfamily\footnotesize,
  breakatwhitespace=false,         
  breaklines=true,                 
  captionpos=b,                    
  keepspaces=true,                 
  numbers=left,                    
  numbersep=5pt,                  
  showspaces=false,                
  showstringspaces=false,
  showtabs=false,                  
  tabsize=2
}
\lstset{style=codestyle}

% Macros
\newcommand{\kt}{Knowledge Tracing}
\newcommand{\kc}{Knowledge Component}
\newcommand{\kcs}{Knowledge Components}
\newcommand{\reals}{\mathbb{R}}
\newcommand{\ece}{\mathrm{ECE}}
\newcommand{\brier}{\mathrm{Brier}}
\newcommand{\unc}{\mathrm{UNC}}
\newcommand{\rel}{\mathrm{REL}}
\newcommand{\res}{\mathrm{RES}}
\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}

% For compatibility with tables that use tableorg
\let\tableorg\table
\let\endtableorg\endtable

\begin{document}

\title{\sffamily\bfseries Reproducible Sparse-Concept and Calibration Diagnostics for Knowledge Tracing\\
{\large A Protocol and Diagnostic Study}}

\author{
\sffamily Khanh-Trinh Nguyen\textsuperscript{1}, Tuan Dao Minh\textsuperscript{1}, Duong Nguyen Tien\textsuperscript{1}, \\
\sffamily Van-Hau Nguyen\textsuperscript{*,1}, Nguyen Chi Thanh\textsuperscript{2}\\[8pt]
\normalsize \textsuperscript{1}\textit{Hung Yen University of Technology and Education}, Hung Yen, Vietnam\\
\normalsize Email: \{trinhnk, tuanymc, duongnt, haunv\}@utehy.edu.vn\\[4pt]
\normalsize \textsuperscript{2}\textit{Academy of Military Science and Technology}, Ha Noi, Vietnam\\
\normalsize Email: thanhnc@ioit.ai.vn\\[8pt]
\normalsize \textsuperscript{*}Corresponding author: haunv@utehy.edu.vn
}
\date{}

\maketitle

\begin{abstract}
Knowledge Tracing (KT) models are typically evaluated with aggregate predictive metrics such as the area under the ROC curve (AUC) and accuracy. Such metrics can mask how a model behaves on sparse and cold-start Knowledge Components (KCs)---concepts with limited or no training-fold evidence---even though reliable probability estimates on these concepts are precisely what educational decision support depends on. Rather than proposing a new model, we treat KT evaluation as a reproducibility and diagnostic problem. We present a reproducible diagnostic protocol that extends standard KT evaluation. It combines learner-based and temporal splits, train-only KC-frequency stratification with an explicit strict cold-start group, and sample-size-aware reporting. Furthermore, it introduces per-stratum calibration analysis using the Expected Calibration Error (ECE), Brier-score decomposition, and reliability diagrams, supported by an eight-channel leakage and predictive-sanity audit. We instantiate the protocol on three datasets---ASSISTments 2012, Junyi Academy, and XES3G5M---with IRT (Item Response Theory), DKT (Deep Knowledge Tracing), and SimpleKT baselines. Under our experimental conditions, on datasets with populated sparse strata, calibration error tends to increase from dense to sparse concepts while aggregate AUC may obscure these stratum-level reliability differences; the predictive-sanity channel additionally flagged a prediction--label misalignment that aggregate metrics alone would not reveal. Sparse-concept behavior varies substantially across datasets, which motivates dataset-local diagnostics rather than reliance on a single aggregate score. We provide an anonymized reproducibility artifact for peer review and will release the public version upon acceptance.
\end{abstract}

\vspace{10pt}
\noindent\textbf{Keywords:} Knowledge Tracing; Educational Data Mining; Calibration; Sparse Concepts; Cold-start Knowledge Components; Reproducibility

\input{sections/01_introduction}
\input{sections/02_related_work}
\input{sections/03_protocol}
\input{sections/04_experiments}
\input{sections/05_discussion_limitations}
\input{sections/06_conclusion}

\section*{Artifact Availability \& Reproducibility Checklist}
\begin{sloppypar}
The anonymized code, data preprocessing scripts, and reproducible evaluation pipelines are available for peer review at \url{https://anonymous.4open.science/r/Sparse-Concept-and-Calibration-6E5B/}. The public repository will be released upon acceptance. Consistent with the design principles of Section~\ref{sec:protocol}, the artifact includes the following standardized components:
\end{sloppypar}
\begin{itemize}
  \item \textbf{Dataset Preprocessing:} Scripts that normalize raw data into the schema defined in Section~\ref{sec:protocol}.
  \item \textbf{Split Construction:} Reproducible partition logs and multi-seed split constructors.
  \item \textbf{Prediction-level Exports:} CSV files containing baseline predicted probabilities and ground-truth labels.
  \item \textbf{KC-Strata Assignment:} Code mapping training-fold frequencies to strata buckets.
  \item \textbf{Calibration Verification:} Scripts computing Expected Calibration Error (ECE).
  \item \textbf{Brier Decomposition:} Implementation of Uncertainty, Reliability, and Resolution equations.
  \item \textbf{Reliability Diagrams:} Plotting utilities for multi-strata calibration curves.
  \item \textbf{Leakage \& Predictive Sanity Audit:} Automated logs verifying channels L1--L8.
  \item \textbf{LaTeX Table Exporters:} Python exporters generating booktabs tables.
  \item \textbf{One-Command Reproduction:} Unified script ensuring machine-independent reproducibility.
\end{itemize}

\appendix
\input{appendix/appendix_a_sensitivity}

\section*{Declaration of Generative AI Software Tools in the Writing Process}
During the preparation of this work, the authors used ChatGPT, Claude and Google Antigravity to support language polishing, LaTeX formatting, consistency checking, and preparation of reproducibility prompts in the Abstract, Introduction, Methods, Results, Discussion, and Appendix sections. These tools were not used to generate, fabricate, or alter experimental results. After using these tools, the authors reviewed and edited the content as needed and take full responsibility for the content of the publication.

\section*{Acknowledgements}
This work was supported by the authors' respective institutions.

\section*{ORCID}
\noindent Khanh-Trinh Nguyen: 0009-0004-3739-0484\\
\noindent Tuan Dao Minh: 0009-0009-1641-6813\\
\noindent Duong Nguyen Tien: 0009-0007-1104-7129\\
\noindent Van-Hau Nguyen: 0000-0002-3256-5626\\
\noindent Nguyen Chi Thanh: 0000-0003-4335-7002.

% \bibliographystyle{acmtrans}
% \bibliography{references}
\input{references.bbl}
\end{document}
